Fast Regression of the Tritium Breeding Ratio in Tokamak Fusion Reactors

### 2020-05-21 paper/methodology.tex
add full methodology section from the report

In the decoupled sampling approach that is further described
in~\cref{sec:supervised}, we first gather a sufficiently large
training set of samples $\mathcal{T}=\left\{\left( x^{(i)},f\left(x^{(i)}\right) \right)\right\}_{i=1}^N$
to describe the behaviour of $f(x)$ across its domain.
Depending on a specific model family and appropriate choice of its
hyperparameters, surrogate models $\hat{f}(x)$ are trained to minimise
the empirical risk~$R_{\text{emp.}}$ with respect to $\mathcal{T}$ and a model-specific
loss function $\mathcal{L}$, where the empirical risk is defined as
$R_{\text{emp.}}(\hat{f}\mid\mathcal{T},\mathcal{L})
	=\frac{1}{N}\sum_{i=1}^N
	\mathcal{L}\left(\hat{f}(x^{(i)}),f(x^{(i)})\right)$.

The adaptive sampling approach that we characterise in~\cref{sec:adaptive} can be viewed as a more general problem.
Rather than fixing the training set $\mathcal{T}$ for the entire duration of
training, multiple sets $\{\mathcal{T}_k\}_{k=0}^K$ are used, such that
$\mathcal{T}_{k-1}\subset\mathcal{T}_k$ for all $k>1$. The first set
$\mathcal{T}_0$ is initialised randomly to provide a \textit{burn-in}, and is
repeatedly extended in epochs, whereby each epoch trains a new surrogate~$\hat{f}_k(x)$ on
$\mathcal{T}_k$ using the supervised learning procedure, evaluates its
performance, and forms a new set $\mathcal{T}_{k+1}$ by adding more samples to
$\mathcal{T}_k$. This permits the learning algorithm to condition the selection
of new samples in~$\mathcal{T}_{k+1}$ on the evaluation results
of~$\hat{f}_k(x)$ in order to maximise improvement of
surrogate performance over complex regions within the feature space.

@@

\Cref{tbl:metrics} provides an exhaustive listing and description of metrics recorded
in our experiments. For regression performance analysis, we include a selection
of absolute metrics to assess the approximation capability of surrogates, and set
practical bounds on the expected uncertainty of their predictions. In addition, we also track
relative measures that are better-suited for comparison between this work and others, as
they maintain invariance with respect to the selected domain and image space.
For complexity analysis, surrogates are assessed in terms of wall
time\footnote{Real time elapsed during computation measured by a
chronometer, here by means of the Python~\texttt{time} package.}
elapsed during training and prediction. This is motivated by common practical use
cases of our work, where models are trained and used as drop-in replacements for the
expensive MC TBR model. Since training set sizes remain to be determined, all times are
reported per a single datapoint. Even though some surrogates support acceleration
by means of parallelisation, sequential processing of samples was ensured to
achieve comparability between considered
models.\footnote{The only exception to this are artificial neural networks,
	which require considerable amount of processing power for training on conventional CPU architectures.}

@@

In our experiments, we evaluate and compare surrogates in effort to
optimise against metrics described in~\cref{sec:metrics}. To attain meaningful
and practically usable results, we require a sufficiently large and diverse pool
of surrogate families to review. This is described by the listing in~\cref{tbl:surrogates}. The presented selection of models includes basic
techniques suitable for linear regression enhanced by the kernel trick or dimension
lifting, methods driven by decision trees, instance-based learning models,
ensemble regressors, randomised algorithms, artificial neural networks and mathematical approaches
developed specifically for the purposes of surrogate modelling. For each of
these families, a state-of-the-art implementation was selected and adapted to
operate with TBR samples.

@@

While some of the presented model families are clearly determined by an explicit
choice of a learning algorithm, others may vary considerably depending on
hyperparameter configuration. A good example of the latter group are artificial neural
networks, which in addition to conventional hyperparameters enable to control network
architecture through selection from various parametric graph templates
(illustrated in~\cref{fig:nn-archs}). This allows us to realise a simplistic
network architecture search during the hyperparameter tuning procedure.

\begin{figure}[h]
	\centering
	\begin{subfigure}[b]{0.25\textwidth}
		\centering
		{\scriptsize \incfigscale{0.6}{1h3f}}
		\caption{$\text{1h3f}(W)$}
	\end{subfigure}\hfill%
	\begin{subfigure}[b]{0.25\textwidth}
		\centering
		{\scriptsize \incfigscale{0.6}{kf}}
		\caption{$\text{df}(D,W)$}
	\end{subfigure}\hfill%
	\begin{subfigure}[b]{0.25\textwidth}
		\centering
		{\scriptsize \incfigscale{0.6}{3pyramid}}
		\caption{$\text{3pyramid}(W)$}
	\end{subfigure}\hfill%
	\begin{subfigure}[b]{0.25\textwidth}
		\centering
		{\scriptsize \incfigscale{0.6}{5diam}}
		\caption{$\text{5diamond}(W)$}
	\end{subfigure}

@@

The aim of the initial experiments is to use a relatively small subset of
collected TBR samples to determine hyperparameters of considered surrogates.
Since this process requires learning the behaviour of an unknown, possibly
expensive mapping -- here a function that assigns cross-validated metrics to a
point in the hyperparameter domain -- it in many aspects mirrors
the primary task of this work with the notable extension of added utility
to optimise. In order to avoid undesirable exponential slowdown in exhaustive
searches of a possibly high-dimensional parameter space, Bayesian
optimisation~\cite{movckus1975bayesian} is employed as a standard hyperparameter tuning algorithm. We set
its objective to maximise $R^2$ and perform
1000~iterations.\footnote{Hyperparameter tuning of each surrogate family was
	terminated after 2~days. Instances that reached this limit may be identified
	in~\cref{tbl:exp1-detailed-results,tbl:exp2-detailed-results} in the
	Appendix.}

@@

\begin{wrapfigure}[25]{r}{0.55\textwidth}
	\centering
	\vspace{-3ex}
	\includegraphics[width=0.7\textwidth]{fig4_qassplan.png}
	\caption{Schematic of QASS algorithm}
	\label{fig:qassplan}
\end{wrapfigure}

@@

We designed the Quality-Adaptive Surrogate Sampling algorithm (QASS,
\cref{fig:qassplan}) to iteratively increment the training/test set with sample
points which maximise surrogate error and minimise a crowding distance metric
(CDM)~\cite{Solonen2012} in feature space. On each iteration following an initial training of the surrogate on $N$ uniformly random samples, the surrogate was trained and absolute error calculated. MCMC was then performed on the error function generated by performing nearest-neighbor interpolation on these test error points. The resultant samples were culled by $50\%$ according to the CDM, and then the $n$ highest-error candidates were selected for reintegration with the training/test set, beginning another training epoch. Validation was also performed during each iteration on independent, uniformly-random sample sets.

>>>>

In the decoupled sampling approach that is further described
in Section~\ref{sec:supervised}, we first gather a sufficiently large
training set of samples $\mathcal{T}=\left\{\left( x^{(i)},f\left(x^{(i)}\right) \right)\right\}_{i=1}^N$
to describe the behaviour of $f(x)$ across its domain.
Depending on a specific model family and appropriate choice of its
hyperparameters, surrogate models $\hat{f}(x)$ are trained to minimise
the empirical risk~$R_{\text{emp.}}$ with respect to $\mathcal{T}$ and a model-specific
loss function $\mathcal{L}$, where the empirical risk is defined as
$R_{\text{emp.}}(\hat{f}\mid\mathcal{T},\mathcal{L})
	=\frac{1}{N}\sum_{i=1}^N
	\mathcal{L}\left(\hat{f}(x^{(i)}),f(x^{(i)})\right)$.

The adaptive sampling approach that we characterise in Section~\ref{sec:adaptive} can be viewed as a more general problem.
Rather than fixing the training set $\mathcal{T}$ for the entire duration of
training, multiple sets $\{\mathcal{T}_k\}_{k=0}^K$ are used, such that
$\mathcal{T}_{k-1}\subset\mathcal{T}_k$ for all $k>1$. The first set
$\mathcal{T}_0$ is initialised randomly to provide a \textit{burn-in}, and is
repeatedly extended in epochs, whereby each epoch trains a new surrogate~$\hat{f}_k(x)$ on
$\mathcal{T}_k$ using the supervised learning procedure, evaluates its
performance, and forms a new set $\mathcal{T}_{k+1}$ by adding more samples to
$\mathcal{T}_k$. This permits the learning algorithm to condition the selection
of new samples in~$\mathcal{T}_{k+1}$ on the evaluation results
of~$\hat{f}_k(x)$ in order to maximise improvement of
surrogate performance over complex regions within the feature space.

@@

Table~\ref{tbl:metrics} provides an exhaustive listing and description of metrics recorded
in our experiments. For regression performance analysis, we include a selection
of absolute metrics to assess the approximation capability of surrogates, and set
practical bounds on the expected uncertainty of their predictions. In addition, we also track
relative measures that are better-suited for comparison between this work and others, as
they maintain invariance with respect to the selected domain and image space.
For complexity analysis, surrogates are assessed in terms of wall
time\footnote{Real time elapsed during computation measured by a
chronometer, here by means of the Python~\texttt{time} package.}
elapsed during training and prediction. This is motivated by common practical use
cases of our work, where models are trained and used as drop-in replacements for the
expensive MC TBR model. Since training set sizes remain to be determined, all times are
reported per a single datapoint. Even though some surrogates support acceleration
by means of parallelisation, sequential processing of samples was ensured to
achieve comparability between considered
models.\footnote{The only exception to this are artificial neural networks,
	which require considerable amount of processing power for training on conventional CPU architectures.}

@@

In our experiments, we evaluate and compare surrogates in effort to
optimise against metrics described in Section~\ref{sec:metrics}. To attain meaningful
and practically usable results, we require a sufficiently large and diverse pool
of surrogate families to review. This is described by the listing in Table~\ref{tbl:surrogates}. The presented selection of models includes basic
techniques suitable for linear regression enhanced by the kernel trick or dimension
lifting, methods driven by decision trees, instance-based learning models,
ensemble regressors, randomised algorithms, artificial neural networks and mathematical approaches
developed specifically for the purposes of surrogate modelling. For each of
these families, a state-of-the-art implementation was selected and adapted to
operate with TBR samples.

@@

While some of the presented model families are clearly determined by an explicit
choice of a learning algorithm, others may vary considerably depending on
hyperparameter configuration. A good example of the latter group are artificial neural
networks, which in addition to conventional hyperparameters enable to control network
architecture through selection from various parametric graph templates
(illustrated in Figure~\ref{fig:nn-archs}). This allows us to realise a simplistic
network architecture search during the hyperparameter tuning procedure.

\begin{figure}[h]
	\centering
	Placeholder % TODO
%	\begin{subfigure}[b]{0.25\textwidth}
%		\centering
%		{\scriptsize \incfigscale{0.6}{1h3f}}
%		\caption{$\text{1h3f}(W)$}
%	\end{subfigure}\hfill%
%	\begin{subfigure}[b]{0.25\textwidth}
%		\centering
%		{\scriptsize \incfigscale{0.6}{kf}}
%		\caption{$\text{df}(D,W)$}
%	\end{subfigure}\hfill%
%	\begin{subfigure}[b]{0.25\textwidth}
%		\centering
%		{\scriptsize \incfigscale{0.6}{3pyramid}}
%		\caption{$\text{3pyramid}(W)$}
%	\end{subfigure}\hfill%
%	\begin{subfigure}[b]{0.25\textwidth}
%		\centering
%		{\scriptsize \incfigscale{0.6}{5diam}}
%		\caption{$\text{5diamond}(W)$}
%	\end{subfigure}

@@

The aim of the initial experiments is to use a relatively small subset of
collected TBR samples to determine hyperparameters of considered surrogates.
Since this process requires learning the behaviour of an unknown, possibly
expensive mapping -- here a function that assigns cross-validated metrics to a
point in the hyperparameter domain -- it in many aspects mirrors
the primary task of this work with the notable extension of added utility
to optimise. In order to avoid undesirable exponential slowdown in exhaustive
searches of a possibly high-dimensional parameter space, Bayesian
optimisation~\cite{movckus1975bayesian} is employed as a standard hyperparameter tuning algorithm. We set
its objective to maximise $R^2$ and perform
1000~iterations.\footnote{Hyperparameter tuning of each surrogate family was
	terminated after 2~days. Instances that reached this limit may be identified
	in Table~\ref{tbl:exp1-detailed-results,tbl:exp2-detailed-results} in the
	Appendix.}

@@

\begin{figure}[h]
	\centering
	Placeholder % TODO: \includegraphics[width=0.7\textwidth]{fig4_qassplan.png}
	\caption{Schematic of QASS algorithm}
	\label{fig:qassplan}
\end{figure}

@@

We designed the Quality-Adaptive Surrogate Sampling algorithm (QASS,
Figure~\ref{fig:qassplan}) to iteratively increment the training/test set with sample
points which maximise surrogate error and minimise a crowding distance metric
(CDM)~\cite{Solonen2012} in feature space. On each iteration following an initial training of the surrogate on $N$ uniformly random samples, the surrogate was trained and absolute error calculated. MCMC was then performed on the error function generated by performing nearest-neighbor interpolation on these test error points. The resultant samples were culled by $50\%$ according to the CDM, and then the $n$ highest-error candidates were selected for reintegration with the training/test set, beginning another training epoch. Validation was also performed during each iteration on independent, uniformly-random sample sets.
